---
title: "Upgrade R Without Losing Your Packages"
author: "Data Science Riot!"
date: "May 23, 2017"
output: md_document
---

Since the first publication of this post, a couple of packages have emerged to automate this process. The installr package for Windows and the updateR package for OS X are particularly good. However, this continues to be a popular post, so I have decided to keep it up. This work-flow is short, sweet, and cross-platform.

1. Before you upgrade, build a temp file with all of your old packages.

```{r}
setwd("~/.Rbackup/")
tmp <- installed.packages()
installedpkgs <- as.vector(tmp[is.na(tmp[,"Priority"]), 1])
# save(installedpkgs, file="installed_old.rda")
```

2. Install the new version of R and let it do it’s thing.

3. Once you’ve got the new version up and running, reload the saved packages and re-install them from CRAN.

```{r}
setwd("~/.Rbackup/")
load("installed_old.rda")
tmp <- installed.packages()
installedpkgs.new <- as.vector(tmp[is.na(tmp[,"Priority"]), 1])
missing <- setdiff(installedpkgs, installedpkgs.new)
install.packages(missing)
update.packages()
```

If you had any packages from BioConductor, you will need to reload those as well.

```{r}
load("installed_old.rda")
tmp <- installed.packages()
installedpkgs.new <- as.vector(tmp[is.na(tmp[,"Priority"]), 1])
missing <- setdiff(installedpkgs, installedpkgs.new)
BiocManager::install(missing)
```

All done, now you can get back to cracking out R code. This method helped me save a lot of time, hope someone else finds it useful!

The following packages were installed from GitHub/Bitbucket manually.

```{r}
devtools::install_github("stephenturner/annotables")
devtools::install_github("glmgen/genlasso")
devtools::install_github("statsmaths/glmgen", subdir="R_pkg/glmgen")
devtools::install_github("YunlongJiao/kernrank")
devtools::install_github("jpvert/kmr")
devtools::install_github("jpvert/kmr4toxicogenetics")
devtools::install_bitbucket("lwaldron/survhd@default")
devtools::install_github("jtleek/tspreg")
devtools::install_github("mcveanlab/TreeWAS/TreeWAS")
```
